---
title: "3818 R Homework 1"
author: "*** STUDENT NAME HERE ***"
date: "8/20/2020"
output: html_document
---

R Markdown: https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html

## Question 1 
Go through introduction_to_R.pdf. Install R and R studio. In the context of the
introduction_to_R.pdf, what is the value of nonsense?

*Answer:* 


## Question 2 
Load the dataset on household and neighbor characteristics found at https://mattbutner.github.io/data/housing_df.csv as done in the introduction_to_R.pdf document. Look in the upper right panel of R studio, how many variables and how many observations are in this data set?

```{r Question 2}
# Code for Question 2 Here

```

*Answer:* 
For question 2, the standard deviation is 10.54

## Question 3 
The variable descriptions for this dataset can be found on canvas. Look at the
housing_df_info.docx. Which of the variables are quantitative, which of the variables are categorical?

*Answer:* 


## Question 4 
Use the function mean() to find the average median value of owner-occupied homes in
$1000’s. Report the R code and number to two digits.

```{r Question 4}
# Code for Question 4 Here


```

*Answer:* 


## Question 5 
Use the function sd() to calculate the sample standard deviation median value of owner-
occupied homes in $1000’s. Report the R code and number to two digits.

```{r Question 5}
# Code for Question 5 Here


```

*Answer:* The standard deviation of the median housing value is $9,197.10.

## Question 6 
Using the functions sqrt(), sum(), ^2, length(), and mean(), but not var() or sd(), calculate
the sample standard deviation median value of owner-occupied homes in $1000’s. Report the R code and number to two digits.

```{r Question 6}
# Code for Question 6 Here


```

*Answer:*

## Question 7
Use the hist() function, create a histogram of NOX pollution. With this distribution is the
mean or median a better measure of central tendency?

```{r Question 7}
# Code for Question 7 Here


```

*Answer:* 