---
title: "3818 R Homework 5"
author: '*** Student Name ***'
date: "8/20/2020"
output: html_document
---



## Question 1 
For this exercise we will run a regression using Swiss demographic data from around 1888. The sample is a cross-section of French speaking counties in Switzerland.
This data come with the R package datasets. The first step is to load the package into your
current environment by typing the command library(datasets) in to the R console. 1 This loads a
number of datasets including one called swiss. Type help(swiss) in the console for additional
details. The basic variable definitions are as follows:

A data frame with 47 observations on 6 variables, each of which is in percent, i.e., in [0, 100].

[,1]	Fertility	Ig, ‘common standardized fertility measure’
[,2]	Agriculture	% of males involved in agriculture as occupation
[,3]	Examination	% draftees receiving highest mark on army examination
[,4]	Education	% education beyond primary school for draftees.
[,5]	Catholic	% ‘catholic’ (as opposed to ‘protestant’).
[,6]	Infant.Mortality	live births who live less than 1 year.

```{r Load 1888 Swiss demographic data}
help(swiss, package="datasets")

data(swiss, package="datasets")

head(swiss)
```


Use the summary() command to report the mean and median for the variables Fertility,
Education, and Catholic.


## Question 2 
We want to estimate the expected Fertility level in a Swiss county conditional on the county’s education level. We assume the relationship is linear. So, we are interested in estimating $\alpha$ and $\beta$ in \[ 
	\text{Fertility}_c = \alpha + \beta \cdot \text{Education}_c + \epsilon_c.
\] If we use Ordinary Least Squares to estimate and we have the following formulas:  \[ 
	\hat{\beta} = r_{x,y} \frac{s_y}{s_x}
\] \[ 
	\hat{\alpha} = \bar{y} - \hat{\beta} \bar{x},
\]
where $y$ is the left hand side variable, $x$ is the right hand side variable, and the bar $\bar{ }$  denotes the mean, $s$ is the standard deviation, and $r_{x,y}$ is the correlation between $x$ and $y$.
- Find the correlation between Education and Fertility using the `cor()` function, as well as the sample standard deviation for each variable using the `sd()` function. Report these numbers.
- Use the `cor()` and `sd()` function to get an estimate for $\beta$ in the equation relating Fertility toEducation. Keep this value stored in a scalar called beta_hat. Report this number.
- Now use the estimate `beta_hat`, along with the function `mean()` to get an estimate for
alpha. Keep this value stored as a scalar called `alpha_hat`. Report this number.

```{r Question 2}
# Code for Question 2 goes here

```

*Answer*:


## Question 3 
Use alpha_hat and beta_hat to predict the average fertility rate in a county where 40% of the population is educated.

```{r Question 3}
# Code for Question 3 goes here

```

*Answer*:



## Question 4 
Plot the relationship between Fertility and Education using the `plot()` function with Education on the horizontal axis. Make sure to label your axis!

```{r Question 4}
# Code for Question 4 goes here

```

*Answer*:




## Question 5 
Now estimate the model the model relating Fertility Rate to Education using the `lm()` function in R’s base code. Typically, if you want to estimate you use the syntax `lm(yvar ~ xvar, data= dataframe)`.
- Store the estimation results as follows `model_1 <- lm(...)`. This list should include a number of details include the estimated parameters, the coefficient of determination (r-squared), all of the residuals from the model, and more.
- Use the command `summary(model_1)` to report the summary of the ordinary least squares estimation and paste the results in the word document. Do you have the same estimates for and from Question 2?
-. What is the R-squared from this regression? Interpret it in a meaningful way.

```{r Question 5}
# Code for Question 5 goes here

```

*Answer*:




## Question 6 
For each one of the estimated parameters reported in Question 5:
- Interpret the coefficient in a meaningful way.
- Report the results from testing the null hypothesis that the true parameter value is zero.

*Answer*:


## Question 7 
Recreate the figure in Question 4, and then add the line of best fit using the `abline()` function
with the coefficients from `model_1`, `model_1$coefficients`.

```{r Question 7}
# Code for Question 7 goes here

```

*Answer*:


## Question 8 
Plot Education with the residuals associated with the model, `model_1$residuals`. Do the residuals show any pattern?

```{r Question 8}
# Code for Question 8 goes here

```

*Answer*:



## Question 9 
Use the `mean()` command to show that the average of the residuals associated with model_1 is zero. 

```{r Question 9}
# Code for Question 9 goes here

```






