---
title: "R Day 2"
output:
  html_document:
    theme: spacelab
    highlight: tango
---

# Binomial Distribution

$$X \sim B(n = 5, p= 0.6)$$
Calculate $P(X=1)$, $P(X = 3)$, and $P(X >= 4)$ by hand. The function `choose(n,k)` will help.

```{r Probability by Hand}
p <- 0.6

# Calculate P(X = 1)
choose(5,1) * p^1 * (1-p)^4

# Calculate P(X = 3)
choose(5,3) * p^3 * (1-p)^2

# Calculate P(X >= 4)
(choose(5,4) * p^4 * (1-p)^1) + (choose(5,5) * p^5 * (1-p)^0)

```

R has a suite of distribution functions that will make this way easier. For each distribution you will have a corresponding set of for functions: `rxxxx`, `pxxxx`, `qxxxx`, and `dxxxx`, but we will only use the first 3 in this class

```{r Help}
help("distributions")
help("Binomial")
help("Normal")
```

# Calculating Probability of Binomial Distribution

Calculate $P(X=1)$, $P(X = 3)$, and $P(X= 4)$ using the `pbinom` function

```{r}
# P(X = 1)
dbinom(x = 1, size = 5, prob = 0.6)
# P(X <= 1)
pbinom(q = 1, size = 5, prob = 0.6)
# P(X = 0) + P(X = 1)
dbinom(x = 0, size = 5, prob = 0.6) + dbinom(x= 1, size = 5, prob = 0.6)

# P(X = 3)
dbinom(x = 3, size = 5, prob = 0.6)

# P(X >= 4)
# P(X = 4) + P(X = 5)
dbinom(x = 4, size = 5, prob = 0.6) + dbinom(x= 5, size = 5, prob = 0.6)

# 1 - P(X <= 3)
1 - pbinom(q = 3, size = 5, prob = 0.6)

```



# Generating Random Variables from Binomial Distribution

Read through the Arguments of Binomial distribution, for the above example, lets try calculating 1 observation of $X$.

```{r Drawing from Binomial Distribution}
rbinom(n = 100, size = 5, prob = 0.6)
```


Let's plot a histogram of the sample distribution of 10 observations with the argument `probability = TRUE` and `breaks = -0.5:0.5`. Do the sample probabilities match the calculated probabilities above?

```{r}
sample <- rbinom(n = 10, size = 5, prob = 0.6)
print(sample)
# Plot Histogram
hist(sample, breaks= -0.5:5.5, probability = TRUE)
```

Repeat the exercise below and increase the number of obsevations. What happens to the sample probabilities as n gets larger?

```{r}
large_sample <- rbinom(n = 10000, size = 5, prob = 0.6)
hist(large_sample, breaks= -0.5:5.5, probability = TRUE)
```

** PRACTICE **

Try the histogram for different values of size and prob. What happens to the distribution?

$$ \mu = np $$

$$ \sigma = \sqrt(np(1-p))$$

```{r Different values of size}
large_sample <- rbinom(n = 10000, size = 1000, prob = 0.6)
hist(large_sample, breaks= -0.5:1000.5, probability = TRUE)
```


```{r Different values of Prob}
large_sample <- rbinom(n = 10000, size = 5, prob = 0.3)
hist(large_sample, breaks= -0.5:5.5, probability = TRUE)
```





# Normal Distribution

Create a histogram of normal distributions using `mean = 0` and `sd = 2`. Compare the Binomial Distribution and the Normal Distribution

```{r}
# help("Normal")
norm_sample <- rnorm(n = 10000, mean = 1000, sd = 40) 

hist(norm_sample, probability = TRUE)
```




