---
title: "t-test, Size, Power, Two-mean t-test"
author: "Kyle F. Butts"
date: "10/28/2020"
output: html_document
---

# Size and Power

## Size

Let's see how often we reject the null when the null is the true population mean (i.e. size of our test). We will use the level of confidence $\alpha = 0.05$.

$$ H_0: \mu = 10$$ 
$$ H_A: \mu \neq 10 $$



```{r title="Size"}
n_sims <- 10000

# Can use normal z-score
n <- 50
alpha <- 0.10

sample_means <- c()
sample_rejects <- c()

for(i in 1:n_sims){
	
	draws <- rnorm(n, mean = 10, sd = 4)
	mean <- mean(draws)

	# Standardize Mean
	z_score <- (mean - 10)/(4/sqrt(n))
	
	# p-value = 2 * P(z <= - |z-score|)
	z_score <- -abs(z_score)
	p_value <- 2 * pnorm(z_score)
	
	# reject if p_value < alpha
	reject <- p_value < alpha

	# Store results to vector
	sample_rejects <- c(sample_rejects, reject)
	sample_means <- c(sample_means, mean)
}

# hist(sample_means)

# Rejection rate
sum(sample_rejects)/n_sims

# Tails are fatter
# hist(sample_means)

```

## Power

Let's see how often we reject the null when the true population mean is different from our null (i.e. power of our test). We will use the level of confidence $\alpha = 0.05$.
$$ H_0: \mu = 10$$

$$ H_A: \mu \neq 10 $$

To calculate the power, first we need to calculate the rejection region

$$ 
	P(\bar{X} \geq X^*) = .025
$$ 

$$ 
	P(Z \geq \frac{X^* - 10}{4/\sqrt{50}}) = .025
$$ 

$$ 
	\frac{X^* - 10}{4/\sqrt{50}} = 1.96 \implies X^* = 11.11
$$ 

Then, power is the probability of being in the rejection region, given a true population mean, in our case $\mu = 11$. 

$$
	P(\bar{X} \geq 11.11 \vert \mu = 11)
$$

$$
	P(Z \geq \frac{11.11 - 11}{4/\sqrt{50}}) = P(Z \geq 0.1944) = 0.422
$$


```{r title="Size"}
n_sims <- 10000

# Can use normal z-score
n <- 50
alpha <- 0.05
true_mean <- 10.5

sample_means <- c()
sample_rejects <- c()

for(i in 1:n_sims){
	
	draws <- rnorm(n, mean = true_mean, sd = 4)
	mean <- mean(draws)

	# Standardize Mean given the null is true
	z_score <- (mean - 10)/(4/sqrt(n))
	
	# p-value = 2 * P(z <= - |z-score|)
	z_score <- -abs(z_score)
	p_value <- 2*pnorm(z_score)
	
	# reject if p_value < alpha
	reject <- p_value < alpha
	
	# Store results to vector
	sample_rejects <- c(sample_rejects, reject)
	sample_means <- c(sample_means, mean)
}

# Rejection rate
sum(sample_rejects)/n_sims

hist(sample_means)

```


# Two-sided t-test

Class Exercise, fill in the missing code to calculate a two-sided t-test by hand.

$$ 
	H_0: \mu_1 - \mu_2 = 0 
$$
$$
	H_A: \mu_1 - \mu_2 > 0
$$

Our z-stat in this case is 

$$
	Z\text{-stat} = \frac{(\bar{X}_1 - \bar{X}_2) - (\mu_1 - \mu_2)}{\sqrt{\sigma_1/n_1 + \sigma_2/n_2}}
$$

And our p-value is 
$$
	P (Z \geq Z\text{-stat}) 
$$

```{r title="two-sided Z-test"}
set.seed(1)

# Random Sample of group 1 and 2
x1 <- rnorm(50, mean = 10, sd = 5)
x2 <- rnorm(35, mean = 11, sd = 5)

# Calculate means
mean1 <- mean(x1)
mean2 <- mean(x2)
		
# Calculate z-stat
diff_in_means <- mean1 - mean2

z_score <- (mean1 - mean2) / sqrt((25/50) + (25/35))
	
# Calculate p-value
p_value <- 1 - pnorm(z_score)

p_value


```


```{r title="two-sided t-test"}
set.seed(1)

# Random Sample of group 1 and 2
x1 <- rnorm(24, mean = 10, sd = 5)
x2 <- rnorm(15, mean = 11, sd = 5)

# Calculate means
mean1 <- mean(x1)
mean2 <- mean(x2)

var1 <- var(x1)
var2 <- var(x2)
		
# Calculate z-stat
t_score <- (mean1 - mean2) / sqrt((var1/24) + (var2/15))
	
# Calculate p-value
# Note you have to supply t_score and degrees of freedom to tnorm()
# Type ?pt into the terminal to learn how to use this
p_value <- pt(t_score, df = 14)

p_value




```



