---
title: "Regression"
author: "Kyle F. Butts"
date: "11/20/2020"
output: html_document
---

# Load ACS Data

```{r, include = FALSE}
df <- read.csv("~/Desktop/3818/R Days/USA_cars_datasets.csv")

df
```



# t-test for car color

Do red cars and black cars sell for the same price?

$$ H_0: \mu_r - \mu_b = 0$$ 
$$ H_A : \mu_r - \mu_b > 0$$


```{r, include = FALSE}

is_red <- df$color %in% c("red", "blue", "yellow")
is_black <- df$color == "black"

n_red <- sum(is_red)
n_black <- sum(is_black)

# df[rows you want, columns you want]
cars_black <- df[is_black, ]
cars_red <- df[is_red, ]

```


```{r}
# 2 sample Z-test
mu_black <- mean(cars_black$price)
mu_red <- mean(cars_red$price)

sd_black <- sd(cars_black$price)
sd_red <- sd(cars_red$price)

z_score <- (mu_red - mu_black)/sqrt(sd_red^2/n_red + sd_black^2/n_black)

1 - pnorm(z_score)

```

Since the p-value is larger than 0.05, with 95% confidence, I fail to reject that red cars and black cars sell for the same price.


# Regression

```{r}
plot(df$mileage, df$price)
```

```{r}
# Removing Outliers
df <- df[df$mileage < 300000, ]
df <- df[df$price < 60000, ]

plot(df$mileage, df$price)
```


# Regressions

Regression is done by `lm()` function. Specifically the format looks like this:

`reg <- lm(df$y_var ~ df$x_var)`

To see the summary, you can type `summary(reg)`.

```{r}
# price = a + b * mileage

reg <- lm(df$price ~ df$mileage)

reg
summary(reg)

```

```{r}
plot(df$mileage, df$price, col= "red") + abline(reg)


```
























# Simpson's Paradox

```{r Load data}
library(palmerpenguins)
data(penguins)

penguins

```

```{r}
# Scatter plot of Bill Depth (mm) and Body Mass (g)
plot(penguins$bill_depth_mm, penguins$body_mass_g)

cor.test(penguins$body_mass_g, penguins$bill_depth_mm)

# Regression
reg1 <- lm(penguins$body_mass_g ~ penguins$bill_depth_mm)

summary(reg1)

plot(penguins$bill_depth_mm, penguins$body_mass_g) + abline(reg1) 

```



```{r}
# Seperately by penguin species
library(ggplot2)

ggplot(penguins, aes(x = bill_depth_mm, y = body_mass_g)) + 
	geom_point(aes(color = species)) + 
	geom_smooth(method = "lm", aes(color = species))

```
















