---
title: "3818 R Project Example"
author: '*** Student Name ***'
date: "8/21/2020"
output: pdf_document
---

I am looking through the American Consumer Survey for the year 2016 in Colorado. 

```{r}
#load the data into R from your computer: 
# similar for PC's, use the other backslash though to find where your file is located, right click on file in folder and click "get info"
ACS_df <- read.csv("~/Desktop/3818/R Project/example_project/ACS_2016_CO.csv")

```


```{r, include= FALSE}

### CLEANING DATA -- DEALING W/ MISSING OBSERVATIONS

#a lot of variables in this dataset are topcoded because of missing data 
max(ACS_df$FTOTINC)
#we don't want to treat this as a legitimate observation! We can subset our dataset to observations that have good data 

#syntax for subset: subset(data frame, observations to keep, variables to keep)
#the expression "!=" in R means "not equal" 
length(ACS_df$FTOTINC)

# Keep only rows with Total Income that isn't 9999999
ACS_df_use <- ACS_df[ACS_df$FTOTINC != 9999999, ]

length(ACS_df_use$FTOTINC)
```

```{r, include = FALSE}
# Remove NAs from income. This will come up a lot in real life data!!
ACS_df_use <- ACS_df_use[!is.na(ACS_df_use$FTOTINC),]

```

The dataset has `r nrow(ACS_df_use)` observations of individuals in Colorado and for each individual there are `r ncol(ACS_df_use)` variables. These variables cover worker characteristics, familial characteristics, and income. This is a one-time survey, so the data is cross-sectional, i.e. we only see one observation per individual. We will consider total family income in $ and categorical variables for gender, marital status, and education level.

For the total family income, after removing missing observations that were classified as an income of $9,999,999, we have the following summary statistics

- Max: `r max(ACS_df_use$FTOTINC, na.rm= TRUE)`
- Min: `r min(ACS_df_use$FTOTINC, na.rm= TRUE)`
- Mean: `r mean(ACS_df_use$FTOTINC, na.rm= TRUE)`
- Standard Deviation: `r sd(ACS_df_use$FTOTINC, na.rm= TRUE)`


```{r, include= FALSE}

### Recoding Variables

# For any variable you're interested in using, make sure the values match the codebook. We don't want to ever treat a categorical variable as quantitative. For example, the "MARST" variable takes values 1-6 - let's recode this as a dummy look through the documentation of this dataset before using any of your variables 

# Our "MARRIED" dummy will take the value 1 if the person is married and 0 if not 
help("ifelse")

# ifelse(test, yes, no)
ACS_df_use$MARRIED <- ifelse(ACS_df_use$MARST < 4, 1, 0)

# this command inputs 1 if MARST<4, meaning person is married and inputs 0 if MARST is 4 or great, meaning person is not married.

mean(ACS_df_use$MARRIED)
#we can see that about 45% of the people in our sample are married 

# You could make a similary dummy variable for race, such as 0 for white and 1 for other. 

# There are other variables you may want to recode. Such as EDUCD. This variable codes how much education has been obtained but the numerical values don't match the years of schooling. For example 015 means an individual only received up to 3rd grade education. Could think of a few different ways to recode.

```

Among married families, the mean income raises to `r mean(ACS_df_use[ACS_df_use$MARRIED == 1, "FTOTINC"], na.rm= TRUE)` which is larger than the total sample and the standard deviation lowers to `r sd(ACS_df_use[ACS_df_use$MARRIED == 1, "FTOTINC"], na.rm= TRUE)`. The maximum and minimum values don't change.


Now, I turn to look at only people who are active in the labor force and study how their income varies with different characteristics.

```{r, include= FALSE}

### Subsetting Data to Labor Force

# Note: Operator for "is exactly equal to" in a conditional statement: == 
ACS_df_lf <- ACS_df_use[ACS_df_use$LABFORCE == 2, ]

length(ACS_df_lf$LABFORCE)
# we dropped about half the observations 


#now let's drop some variables we don't need 
summary(ACS_df_lf$YEAR)

#we can see that the entire dataset is from 2016, so we don't really need the "YEAR" variable 
ACS_df_lf$YEAR <- NULL 


#we've dropped this variable - let's see what else we might want to drop 
#remember, the summary() function can give us descriptive statistics to see what might be useful 
summary(ACS_df_lf$DATANUM)
summary(ACS_df_lf$STATEFIP)

#looks like DATANUM is 1 for everyone 
#STATEFIP is 8 for everyone 

# Looking in the documentation, this tells us entire sample is from Colorado 
ACS_df_lf$DATANUM <- NULL
ACS_df_lf$STATEFIP <- NULL
```


## Distribution of Income within the Labor Force

Let's look at the distribution of income among the entire Labor Force sample first.

```{r, echo= FALSE}
hist(ACS_df_use$FTOTINC, 
	 main = "Earnings for Colorado Labor Force Participants, 2016", 
	 xlab = "Total Income", col = "steelblue", breaks = 100)
```



The first thing I'm considering is to see if the distribution of earnings differ between men and women. As you can see in the graph below, the distribution looks quite similar between females and male workers in Colorado. 

```{r Female vs. Male Earnings, echo= FALSE}

# What does the distribution of earnings look like? 
ACS_df_lf_men <- ACS_df_lf[ACS_df_lf$SEX==1, ]
ACS_df_lf_women <- ACS_df_lf[ACS_df_lf$SEX==2, ]

# These will be put into your document when you "Knit" it.
hist(ACS_df_lf_men$FTOTINC, 
	 main = "Earnings for Male Colorado Labor Force Participants, 2016", 
	 xlab = "Total Income", col = "steelblue", breaks = 100)
hist(ACS_df_lf_women$FTOTINC, 
	 main = "Earnings for Female Colorado Labor Force Participants, 2016", 
	 xlab = "Total Income", col = "steelblue", breaks = 100)

```


The confidence intervals for mean income for men and women, respectively is:

```{r Confidence Interval Men}

c(mean(ACS_df_lf_men$FTOTINC) - qnorm(.975) * sd(ACS_df_lf_men$FTOTINC)/sqrt(nrow(ACS_df_lf_men)),
	mean(ACS_df_lf_men$FTOTINC) + qnorm(.975) * sd(ACS_df_lf_men$FTOTINC)/sqrt(nrow(ACS_df_lf_men)))

```

```{r Confidence Interval Women}

c(mean(ACS_df_lf_women$FTOTINC) - qnorm(.975) * sd(ACS_df_lf_women$FTOTINC)/sqrt(nrow(ACS_df_lf_women)),
	mean(ACS_df_lf_women$FTOTINC) + qnorm(.975) * sd(ACS_df_lf_women$FTOTINC)/sqrt(nrow(ACS_df_lf_women)))

```

Notice that Women's mean income confidence interval is less than the Men's mean income confidence interval. I'm going to test if female income is lower than men's income on average in our data. From the two-sample t-test below, we can reject with 95% confidence that men and women don't earn the same income on average. 

```{r two-sample t-test}

# You'll see we have two different datasets representing these groups
## Remember, to conduct a hypothesis test, we'll need the mean, std deviation, and sample size of each
t.test(x= ACS_df_lf_men$FTOTINC, y= ACS_df_lf_women$FTOTINC, 
	   alternative= "two.sided",
	   conf.level= 0.95
	  )
```


## Scatter Plot



```{r}

#relationship between age and earnings for college graduates 
#let's subset to people of working age who have a degree and are employed  
ACS_df_grads <- subset(ACS_df_lf, ACS_df_lf$AGE >= 18 & ACS_df_lf$AGE <= 55 & ACS_df_lf$EDUCD >=101 & ACS_df_lf$EMPSTAT ==1) 
length(ACS_df_grads$FTOTINC)
#what does the distribution of earnings look like? 
hist(ACS_df_grads$FTOTINC, main = "Earnings for Colorado Labor Force Participants, 2016"
     ,xlab = "Total Income", col = "steelblue", breaks = 100)

#let's drop outliers 
#how do we make a scatterplot? 


```

```{r}
help("plot")
plot(ACS_df_grads$AGE, ACS_df_grads$FTOTINC, 
		main = "Age and Earnings, CO College Grads, 2016",
		xlab = "Age", ylab = "Total Income", col = "steelblue"
	 ) 

```
